---
title: "Assignments"
output:
  pdf_document:
    toc: no
    toc_depth: '1'
  html_document:
    toc: yes
    toc_float: yes
    collapsed: no
    number_sections: no
    toc_depth: 1
---


This page will contain all the assignments you submit for the class.



### Instructions for all assignments

I want you to submit your assignment as a PDF, so I can keep a record of what the code looked like that day. I also want you to include your answers on your personal GitHub website. This will be good practice for editing your website and it will help you produce something you can keep after the class is over.

1. Download the Assignment1.Rmd file from Canvas. You can use this as a template for writing your answers. It's the same as what you can see on my website in the Assignments tab. Once we're done with this I'll edit the text on the website to include the solutions.

2. On RStudio, open a new R script in RStudio (File > New File > R Script). This is where you can test out your R code. You'll write your R commands and draw plots here.

3. Once you have finalized your code, copy and paste your results into this template (Assignment 1.Rmd). For example, if you produced a plot as the solution to one of the problems, you can copy and paste the R code in R markdown by using the ` ``{r} ``` ` command. Answer the questions in full sentences and Save.

4. Produce a PDF file with your answers. To do this, knit to PDF (use Knit button at the top of RStudio), locate the PDF file in your docs folder (it's in the same folder as the Rproj), and submit that on on Canvas in Assignment 1.

5. Build Website, go to GitHub desktop, commit and push. Now your solutions should be on your website as well.






# Assignment 1

**Collaborators: **


### Problem 1 

Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.

```{r}
#install.packages("datasets")
library(datasets)

USArrests
```

Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?

It is useful to renamed USArrests to dat.us because it is easier to write and it is good practice to rewrite data for yourself so you can create your own data which can be replicated by another person if they use the original data.  


```{r}
dat.us <- USArrests

head(dat.us)
```

### Problem 2

Use this command to make the state names into a new variable called State. 

```{r, eval=TRUE}
dat.us$state <- tolower(rownames(USArrests))
```

This dataset has the state names as row names, so we just want to make them into a new variable. We also make them all lower case, because that will help us draw a map later - the map function requires the states to be lower case.


List the variables contained in the dataset `USArrests`.

```{r}
names(dat.us)
```

**Answer**: The four variables are Murder, Assault, UrbanPop, Rape. 


### Problem 3 

What type of variable (from the DVB chapter) is `Murder`? 

**Answer**: Murder is a quantitative variable. 

What R Type of variable is it?

**Answer**: Murder is numeric. 


### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

**Answer**: The dataset contains the data of murder, assault, rape and urbanpop from all 50 US states. The numbers represent the frequency of arrests for one of the four variables in a state during the time frame that the data was collected.


### Problem 5

Draw a histogram of `Murder` with proper labels and title.

```{r}
hist(dat.us$Murder, xlab= "Murders", ylab="Frequency", main= "Murder in the US", xlim=(c(0, 20)), ylim=(c(0, 13)), col="red", breaks= 10)
```

### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

```{r}
summary(dat.us$Murder)

```

**Answer**: The mean for Murder is 7.788 and the median is 7.250. The mean is the amount that each subject would have if all of the values were added together and evenly distributed. If all 50 states had the same frequency of arrests for murders then it would be 7.788. The median is the middle value where exactly 50% of the values fall either above or below it. In the US, 50% of states have an arrest for murder frequency above 7.250 and the other 50% is below that. The median is highly robust because it is not greatly affected by outliers. The mean is the most common measure of central tendency but it is not robust because it will change based on the skewness of the distribution. A quartile indicates an interval that contains 25% or a quarter of the data.The first quartile for "Murder" is 4.075 which means that 25% of the "Murder" data falls below 4.075 and the 3rd quartile is 11.250 which means that 25% of US states have a frequency of arrests for murder that is higher than 11.250. R gives you the 1st and 3rd quartile because those values are useful in determining the interquartile range (IQR). The IQR is the central half which means that 50% of the data falls within the 1st and 3rd quartile. In a box plot, values 1.5 IQRs above or below the tails are considered outliers.

### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r, echo = TRUE, fig.width = 5, fig.height = 8}
par(mfrow=c(3,1))

hist(dat.us$Murder, xlab= "Murders", ylab="Frequency", main= "Murder in the US", xlim=(c(0, 20)), ylim=(c(0, 13)), col="red", breaks= 10)

hist(dat.us$Assault, xlab= "Assaults", ylab="Frequency", main= "Assault in the US", xlim=(c(0, 350)), ylim=(c(0, 15)), col="blue", breaks=10)

hist(dat.us$Rape, xlab= "Rapes", ylab="Frequency", main= "Rape in the US", xlim=(c(0, 50)), ylim=(c(0, 15)), col="green", breaks=10)

```

What does the command par do, in your own words (you can look this up by asking R `?par`)?

**Answer**: Command par is used to set parameters. The mfrow input allows you to create an array to plot multiple graphs on one window. The command par(mfrow=c(3,1)) allows three graphs to be plotted in three rows. 

What can you learn from plotting the histograms together?

Answer: When the histograms are plotted together it is easier to compare the skewness and spread of each plot. You can see where each histogram has its peaks and outliers. 
  
### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.

Run this code:

```{r, eval= TRUE, fig.width = 7.5, fig.height = 4}

#install.packages("maps") 
#install.packages("ggplot2")

library(maps) 
library(ggplot2) 

ggplot(dat.us, aes(map_id=state, fill=Murder)) + 
  geom_map(map=map_data("state")) + 
  expand_limits(x=map_data("state")$long, y=map_data("state")$lat)
```

What does this code do? Explain what each line is doing.

Answer: The lines library('maps') and library('ggplot2') are pulling from the packages that were installed. The line ggplot(dat.us, aes(map_id=state, fill=Murder)) is creating a ggplot with the USArrests dataset. The plot is set with an aesthetic of a map with the US states. Each state is filled in with its respective murder arrest data. The line geom_map(map=map_data("state")) contains the map coordinates for each US state.The last line  'expand_limits(x=map_data("state")$long, y=map_data("state")$lat)' ensures that the limits of the plot include a single value for all plots. The x and y axis of this plot contains the value of "state" from the map data. The x axis is longitude and the y axis is latitude. Together this code creates a map of the US with each state filled in with its value for murder arrests.The darker blue indicates that the murder arrest frequency is 5 and below and the light blue indicates that it is 15 and above. 


$$\\[2in]$$




# Assignment 2 

(Coming soon)


